---
title: "EDA"
author: "Serena Leung"
date: "4/18/2022"
output: html_document
---

```{r}
library(readr)
NPD <- read_csv("NPD_clean.csv")
head(NPD)  
```
########## Plotting Histograms ################

Age: 
- peak at around 18-20, probably a lot of college students took this survey 
- skewed towards the younger ages 


Score: 
- skewed towards lower scores (less narcissism)

```{r}
library(tidyverse)
library(ggplot2)

ggplot(NPD, aes(x=age)) + 
  geom_histogram(bins = 70) + 
  xlim(1,100)

ggplot(NPD, aes(x=score)) + 
  geom_histogram() + 
  xlim(1, 50)

ggplot(NPD, aes(x = elapse)) + 
  geom_histogram() + 
  xlim(1, 3000)
```

########## Summary Statistics ############
Age:
- Why is there someone that's 509 years old? 
- it said ages 14 and below have been omitted from the dataset

Score:
- how does scoring work? (if answer as narcissist, then give one point)
- max score = 40

```{r}
print("Age")
min(NPD$age)
max(NPD$age)
mean(NPD$age)


print("Score")
min(NPD$score)
max(NPD$score)
mean(NPD$score)
```
```{r}
age_count <- NPD %>% filter(age < 90 & age > 13) %>% select(age) %>% group_by(age) %>% count()

plot(ecdf(age_count$n))

ggplot(age_count, aes(x=age, n)) + 
  geom_point() + 
  xlim(1,100) + 
  geom_segment(aes(x=14, y = 58, xend = 17, yend = 220, color = "segment")) + 
  geom_segment(aes(x=15, y = 58, xend = 18, yend = 220, color = "segment"))



ggplot(age_count, aes(n)) +
  stat_ecdf(geom = "point") + 
  xlim(-1,400) + 
  geom_segment(aes(x=0, y = 0, xend = 30, yend = 0.28, color = "Age 14 - 28"), size = 1.25) + 
  geom_segment(aes(x=45, y = 0.27, xend = 130, yend = 0.57, color = "Age 29 - 45"), size = 1.25) + 
  geom_segment(aes(x=145, y = 0.57, xend = 180, yend = 0.72, color = "Age 46 - 53"), size = 1.25) + 
  geom_segment(aes(x=195, y = 0.72, xend = 250, yend = 0.85, color = "Age 54 - 61"), size = 1.25) +
  geom_segment(aes(x=270, y = 0.85, xend = 310, yend = 0.9, color = "Age 62 - 65"), size = 1.25) + 
  geom_segment(aes(x=330, y = 0.9, xend = 400, yend = 1, color = "Age 66 - 72"), size = 1.25) 
```
Q: Why are we segmenting by slope?

Age Group Segments:  

* Age 12-28 (1)
* Age 29-45 (2)
* Age 46-53 (3)
* Age 54-61 (4)
* Age 62-65 (5)
* Age 66-86 (6)


Score Group Segments: 

* Score 0-5 (1)
* Score 6-14 (2)
* Score 15-25 (3)
* Score 26-32 (4)
* Score 33-40 (5)

```{r}
score_count <- NPD %>% filter(age < 90 & age > 13) %>% select(score) %>% group_by(score) %>% count()

ggplot(score_count, aes(n)) +
  stat_ecdf(geom = "point") + 
  geom_segment(aes(x=0, y = 0, xend = 45, yend = 0.15, color = "Score 1 - 5"), size = 1.25) + 
  geom_segment(aes(x=60, y = 0.15, xend = 120, yend = 0.38, color = "Score 6 - 14"), size = 1.25) + 
  geom_segment(aes(x=135, y = 0.37, xend = 290, yend = 0.65, color = "Score 15 - 25"), size = 1.25) + 
  geom_segment(aes(x=320, y = 0.65, xend = 450, yend = 0.80, color = "Score 26 - 32"), size = 1.25) +
  geom_segment(aes(x=480, y = 0.83, xend = 570, yend = 1, color = "Score 33 - 40"), size = 1.25) 
  
```

############# Mutating to add age group and score group  ###############

```{r}
NPD_group <- NPD %>% mutate(age_group = case_when(
        age >= 12 & age <= 28 ~ 1,
        age >= 29 & age <= 45 ~ 2,
        age >= 46 & age <= 53 ~ 3,
        age >= 54 & age <= 61 ~ 4,
        age >= 62 & age <= 65 ~ 5,
        age >= 66 & age <= 86 ~ 6), 
        
        score_group = case_when(
          score >= 0 & score <= 5 ~ 1,
          score >= 6 & score <= 14 ~ 2,
          score >= 15 & score <= 25 ~ 3,
          score >= 26 & score <= 32 ~ 4,
          score >= 33 & score <= 40 ~ 5))
```

```{r}
NPD_group %>% filter(is.na(age), is.na(score))
```


Trying to create a contingency table with sums and probability

```{r}
score_age <- NPD_group %>% select(age_group, score_group) %>% 
  group_by(age_group, score_group) %>% 
  count() %>%
  pivot_wider(names_from = score_group, values_from = n)    

score_age

row_sums <- rowSums(score_age)
col_sums <- colSums(score_age)[2:6]

row_total <- sum(rowSums(score_age))
col_total <- sum(colSums(score_age)[2:6])

row_sums/row_total
col_sums/col_total 

score_age$age_group<- as.numeric(score_age$age_group)
score_age["row_sum"] <- row_sums
score_age["row_prob"] <- row_sums/row_total
score_age <- rbind(score_age, "col_sum" = col_sums)
score_age <- rbind(score_age, "col_prob" = col_sums/col_total)

score_age

```

######## Gender, Score, Age Contingency Table ############
```{r}
# gender: 1 = M, 2 = F, 3 = Other

NPD_group %>% mutate(gender_score = case_when(
  score_group == 1 & gender == 1 ~ "M1", 
  score_group == 1 & gender == 2 ~ "F1", 
  score_group == 1 & gender == 3 ~ "O1",
  score_group == 2 & gender == 1 ~ "M2",
  score_group == 2 & gender == 2 ~ "F2",
  score_group == 2 & gender == 3 ~ "O2",
  score_group == 3 & gender == 1 ~ "M3",
  score_group == 3 & gender == 2 ~ "F3",
  score_group == 3 & gender == 3 ~ "O3",
  score_group == 4 & gender == 1 ~ "M4",
  score_group == 4 & gender == 2 ~ "F4",
  score_group == 4 & gender == 3 ~ "O4",
  score_group == 5 & gender == 1 ~ "M5",
  score_group == 5 & gender == 2 ~ "F5",
  score_group == 5 & gender == 3 ~ "O5"
)) %>% group_by(gender_score, age_group) %>% 
  select(age_group, gender_score) %>%
  count() %>%
  pivot_wider(names_from = gender_score, values_from = n)  
```


Entropy Formula 

$$H(X) = -\sum^{n}_{i=1} p(x_i) log_2p(x_i)  $$

########### Score Entropy ##############
```{r}
px = col_sums/col_total

-sum(px*log2(px))
```

########### Age Entropy ##############
```{r}
px = row_sums/row_total

-sum(px*log2(px))
```
########## Q1, Age = 1, Score = 1 ##################

```{r}
NPD_group %>% filter(age_group == 1, score_group == 1) %>% select(Q1) %>% group_by(Q1)%>% count()
```

What does it mean if a question has a high entropy? Shouldn't all question equally add to the score?


Q1, Score = 1
```{r}
NPD_group %>% filter(score_group == 1) %>% 
  select(Q1, age_group) %>% 
  group_by(Q1, age_group)%>% 
  count() %>% 
  pivot_wider(names_from = age_group, values_from = n)
```

Q1, Age = 1
```{r}
q1_age1 <- NPD_group %>% filter(age_group == 1) %>% 
  select(Q1, score_group) %>% 
  group_by(Q1, score_group)%>% 
  count() %>% 
  pivot_wider(names_from = score_group, values_from = n)

q1_age1 

row_sums <- rowSums(q1_age1)
col_sums <- colSums(q1_age1)[2:6]

row_total <- sum(rowSums(q1_age1))
col_total <- sum(colSums(q1_age1)[2:6])

paste("H(Score|Q1) for age_group = 1", row_sums/row_total)
paste("H(Score) when age_group = 1", col_sums/col_total)


```

########### Function to Calculate Conditional Entropy ######### 
```{r}

```


############ Mutual Information of Q1 and Q4 [H(Q1) - H(Q1|Q4)] ################
```{r}
## Q1 Q4 contingency table 
q1_q4 <- NPD_group %>% 
  select(Q1, Q4) %>%
  group_by(Q4, Q1) %>% 
  count() %>% pivot_wider(names_from = Q1, values_from = n)

q1_q4

row_sums <- rowSums(q1_q4)
col_sums <- colSums(q1_q4)[2:3]

row_total <- sum(rowSums(q1_q4))
col_total <- sum(colSums(q1_q4)[2:3])

paste("H(Q1|Q4)", row_sums/row_total)
paste("H(Q1)", col_sums/col_total)

p_q1 <- col_sums/col_total
p_q1_q4 <- row_sums/row_total

-sum(p_q1*log2(p_q1)) - -sum(p_q1_q4*log2(p_q1_q4)) ## total for the whole table with no subsetting of score
```


Next Steps: 
- If I want to include score_group, I must subset before calculating H(Q1|Q4)


What do I want to do? (focus on age and score)
- summary statistics (age and score) 
- regular plotting to figure out any obvious trends (hist, geom_point... )
- age: create age groups?
- plot histogram using bins by age groups
- contingency table
- calculate entropy


